Age-related changes in plant growth and traits

### 2015-05-06 results_RAschedules.Rmd
- Results_best new file with most interesting graphs and stories - Thoughts about RA schedules and RV curves mostly developed - Making progress with accessory tissues - Wrote a looping function that works to add se bars - very pleased to have gotten that

@@

@@

@@

Now my current analyses still lack the ability to tease apart the most obvious pattern - because I don't actually want to show that growth and height are correlated, so much as that the asymptoting of growth is correlated with height - i.e. the change in slope of the relationship.

```{r, echo=FALSE, message=FALSE}
yvar <- "GrowthInv"
xvar <- "RA"
x2var <- "PropH"

@@

results$r.squared <- round(results$r.squared,digits=3)
results$p.value <- round(results$p.value,digits=3)

for(spp in names(data)) {
  plot_yvar_vs_xvar(data[[spp]], yvar, xvar, log="y",
                    main=spp, xlab=xvar, ylab=yvar)
  abline(lmfits[[spp]])
}
```

>>>>

plot(ReproInv~RGR, SummaryInd, log="xy", col=col.spp[as.factor(species)],pch=16)
legend("bottomleft",legend=names(labels.spp),col=col.spp,pch=16, cex=.8,bty="n")

@@

  abline(lmfits[[spp]])
}

yvar <- "RGR"
xvar <- "ReproInv"

data <- subset(SummaryInd,RA>0&RGR>.04)
data <- split(data, data$species)
lmfits <- lapply(data, function(x) lm(log(x[[yvar]]) ~  (log(x[[xvar]]))))
results <- summarise_all_fits(lmfits)

@@

Model analyzed is: lm(log([[GrowthInv]]) ~  asin(sqrt([[RA]]))*asin(sqrt([[PropH]]))))

@@

Now my current analyses still lack the ability to tease apart the most obvious pattern - because I don't actually want to show that growth and height are correlated, so much as that the asymptoting of growth is correlated with height - i.e. the change in slope of the relationship.
```{r, echo=FALSE, message=FALSE}
yvar <- "GrowthInv"
xvar <- "RA"
x2var <- "PropH"

@@

results$r.squared <- round(results$r.squared,digits=3)
results$p.value <- round(results$p.value,digits=3)
names(results) <- c("species","r.sq_total","p-total","n","intercept","coef_RA","coef_PropH","coef_interaction","SS_RA","SS_PropH","SS_interaction")
results

yvar <- "GrowthInv"
xvar <- "PropH"

data <- subset(SummaryInd)
data <- split(data, data$species)
lmfits <- lapply(data, function(x) lm(log(x[[yvar]]) ~  asin(sqrt(x[[xvar]]))))
results2 <- summarise_all_fits(lmfits)

results2$r.squared <- round(results2$r.squared,digits=3)
results2$p.value <- round(results2$p.value,digits=3)
names(results2) <- c("species","r.sq.PropH","p.PropH","n","a","b")
results2

data <- subset(SummaryInd)
data <- split(data, data$species)
lmfits <- lapply(data, function(x) lm(log(x[[yvar]]) ~  asin(sqrt(x[[xvar]]))))
results3 <- summarise_all_fits(lmfits)

results3$r.squared <- round(results3$r.squared,digits=3)
results3$p.value <- round(results3$p.value,digits=3)
names(results3) <- c("species","r.sq.RA","p.RA","n","a","b")
results3

results <- merge(results, select(results2, -n, -a,-b), by="species")
results <- merge(results, select(results3, -n, -a,-b), by="species")
results
```

### 2015-05-06 results_accessory_tissues.rmd
- Results_best new file with most interesting graphs and stories - Thoughts about RA schedules and RV curves mostly developed - Making progress with accessory tissues - Wrote a looping function that works to add se bars - very pleased to have gotten that

@@

plot_yvar_vs_xvar(data, yvar, xvar,
                 xlab=xvar, ylab=yvar)
mod <- lm(asin(sqrt(PerPropagule)) ~ species, data)
summary(mod)
```

@@

###Change in accesory costs With respect to RA:  
- both RA and PerAcc are arcsine transformed  
- BAER, COER, GRBU, GRSP, HATE, LEES, PELA, PEPU show a significant decrease in accessory costs with increasing RA (p<0.05)  
- BOLE shows a marginally significant decrease in accessory costs with increasing RA (p ~ 0.10)  
```{r, echo=FALSE, message=FALSE}
yvar <- "PerAcc"
xvar <- "RA"

>>>>

se_function <- function(mean1,se1) {
for(i in 1:14) {
arrows(i,mean1[i]+se1[i],i,mean1[i]-se1[i],code=3,angle=90, length=0.1)
}
}
```

@@

plot_yvar_vs_xvar(data, yvar, xvar,
                 xlab=xvar, ylab=yvar)
mod <- lm(asin(sqrt(PerPropagule)) ~ species-1, data)
summary(mod)
data.frame(
    select(glance(mod), r.squared, p.value),
    n= length(resid(mod)),
    a = coef(mod)[1],
    b = coef(mod)[2]
  )

tmp <- summary(mod)
tmp <- tmp$coefficients
tmp <-as.data.frame(tmp)
tmp <- cbind(names(labels.spp),tmp)
names(tmp) <- c("species","estimate.trans","se.trans","t.value","p.value")
tmp2 <- select(SummarySpp, species, PerPropagule_mean, PerPropagule_se)
tmp <- merge(tmp, tmp2, by="species")
tmp
plot(PerPropagule_mean~species,tmp,cex.axis=.8,pch=16,log="y",ylim=c(.001,.25),main="Percent reproductive investment put into propagules")

mean1 <- tmp$PerPropagule_mean
se1 <- tmp$PerPropagule_se

se_function(mean1,se1)

#for(i in 1:14) {
#arrows(i,tmp$PerPropagule_mean[i]+tmp$PerPropagule_se[i],i,tmp$PerPropagule_mean[i]-tmp$PerPropagule_se[i],code=3,angle=90, length=0.1)}

#barplot(height=test$PerPropagule_mean, names=test$species,cex.axis=.8,ylim=c(0,.25),cex.names=.8)
#xv <- barplot(height=test$PerPropagule_mean, names=test$species,cex.axis=.8,ylim=c(0,.25),cex.names=.8)
#xv

#points((PerPropagule_mean+PerPropagule_se)~species,test,pch=25)
#points((PerPropagule_mean-PerPropagule_se)~species,test,pch=24)

#Tuk <- TukeyHSD(mod)
```

@@

###Change in accesory costs With respect to RA:  
- both RA and PerAcc are arcsine transformed  
- BAER, COER, GRBU, GRSP, HATE, LEES, PELA, PEPU show a significant decrease in accessory costs with increasing RA (p<0.05)  
- BOLE shows a marginally significant decrease in accessory costs with increasing RA (p ~ 0.10)  
```{r, echo=FALSE, message=FALSE}
yvar <- "PerPropagule"
xvar <- "RA"

### 2015-05-06 results_best.Rmd
- Results_best new file with most interesting graphs and stories - Thoughts about RA schedules and RV curves mostly developed - Making progress with accessory tissues - Wrote a looping function that works to add se bars - very pleased to have gotten that

summarise_all_fits_2var <- function(x) {
  tmp <- lapply(x, summarise_fit_2var) %>%
    bind_rows
  data.frame(species=names(x), tmp)
}
```

#Ideas and plots to include in a paper looking at the ways in which reproductive investment impacts growth
* Part of this paper would be to look at how different "types of" species fall out with respect to the various plots - i.e. canopy species (BAER, HATE, PELA), short-lived species (BOLE, PILI, COER), longer-lived understory species 
* Plants that grow more also reproduce more, but plants that allocate more to reproduction have lower RGR (Same paper as RA curves?)
* There is a disinct decline in RGR (or Growth Investment) at the exact "PropH" that plants begin to reproduce (Same paper as RA curves?)

@@

#Ideas and plots to include in RA schedule paper
* I have a dataset with enough resolution across species and ages to show patterns of growth and reproductive investment across ages  
* This includes both RA schedules and RV curves, as well as other potential plots

###RV curves (by proportion max H)
- In a sense, by using proportion max H, these figures are showing the proportion of total growth the plant undergoes before starting reproduction
- You could also plot against "PropMaxWeight", the proportion of the heaviest individual of that species. The two measures look "best" for different species
```{r, echo=FALSE,message=FALSE}
plot((ReproInv)~(PropH), SummaryInd, log="y",col=col.spp[as.factor(species)],pch=16,xlim=c(4e-03,1))
legend("topleft",legend=names(labels.spp),col=col.spp,pch=16, cex=.8,bty="n")

@@

#RA schedules (vs. propH)  
- these curves are particularly good for a number of species such as EPMI, COER, PELA  
- they also show that for PILI that among individuals that reproduce, the "potential" RA decreases with increasing propH  
- when all data points are included, all species except PEPU, PHPH, and PULU show increasing RA with increasing height  
- when RA =0 is excluded, the following species still have a significant positive correlation: BOLE, COER, EPMI, GRBU, GRSP, HATE. LEES and PELA have a p-value between 0.05-0.10   
- PEPU has a negative correlation - matches Mark's suggestion that it is "hanging on"  
- There are species - such as BOLE - which have a much "cleaner" plot against age instead of PropH (or any other size measure). This indicates that these species are programmed to reproduce at a certain "rate" regardless of their size

@@

>>>>

summarise_all_fits_2var <- function(x) {
  tmp <- lapply(x, summarise_fit_2var) %>%
    bind_rows
  data.frame(species=names(x), tmp)
}

xy.error.bars <- function (x,y,xbar,ybar) {
  plot(x,y,pch=16, ylim=c(min(y-ybar),max(y+ybar)),xlim=c(min(x-xbar),max(x+xbar)))
  arrows(x,y-ybar,x,y+ybar,code=3,angle=90, length=0.1)
  arrows(x-xbar,y,x+xbar,y,code=3,angle=90, length=0.1)
}

se_function <- function(mean1,se1) {
for(i in 1:14) {
arrows(i,mean1[i]+se1[i],i,mean1[i]-se1[i],code=3,angle=90, length=0.1)
}
}
```

#Changes in growth and tissue allocation with age and size
* Part of this paper would be to look at how different "types of" species fall out with respect to the various plots - i.e. canopy species (BAER, HATE, PELA), short-lived species (BOLE, PILI, COER), longer-lived understory species

@@

#RA schedule paper
* I have a dataset with enough resolution across species and ages to show patterns of growth and reproductive investment across ages  
* This includes both RA schedules and RV curves  
* Below are curves plotted relative to "PropH", which is a better fit for most species - but other species should be plotted against total plant weight  
* Plants that grow more also reproduce more, but plants that allocate more to reproduction have lower RGR 
* There is a disinct decline in RGR (or Growth Investment) at the exact "PropH" that plants begin to reproduce 
* Plotting reproductive investment (ReproInv) against RGR yields a much lower r^2 - for every single species. Only RA captures the actual trade-off incurred by plants that allocate more to reproduction, showing one way RA schedules are important  
  - As discussed in the review, RA schedule differentiate between greater reproductive output because you are bigger (a positive correlation with plant growth) and greater reproductive output because you are allocating more to reproduction (a negative correlation with plant growth)  
* RV curves continue to be the more relevent metric to determine the actual amount of reproductive output produced by a plant of a given size or height

##RV curves (by proportion max H)
- In a sense, by using proportion max H, these figures are showing the proportion of total growth the plant undergoes before starting reproduction
- The plot across all species of reproductive investment vs. Prop H is much less tight than reproductive investment vs. total weight, but within a species not as obviously different  
- The plot that includes all species shows that *once* a given species has ramped up reproduction, the log-log plot of reproductive investment vs final weight is nearly linear across all species; in other words, to first order, all species allocate the same amount to reproduction relative to their weight.  
- This is presumably the basic allometric scaling relationship discussed by others and wondering if slope I get is same as theirs    
- However, it is important to note that species early in their reproductive trajectory do not follow this pattern  
```{r, echo=FALSE,message=FALSE}
plot((ReproInv)~(PropH), SummaryInd, log="y",col=col.spp[as.factor(species)],pch=16,xlim=c(4e-03,1))
legend("topleft",legend=names(labels.spp),col=col.spp,pch=16, cex=.8,bty="n")

plot(ReproInv~FinalWeight, SummaryInd, log="xy",col=col.spp[as.factor(species)],pch=16)
legend("topleft",legend=names(labels.spp),col=col.spp,pch=16, cex=.8,bty="n")

plot(ReproInv~height, SummaryInd, log="xy",col=col.spp[as.factor(species)],pch=16)
legend("topleft",legend=names(labels.spp),col=col.spp,pch=16, cex=.8,bty="n")

@@

##RA schedules (vs. propH)  
- RA schedules still look "best" plotted against age; not shown here  
- these curves are particularly good for a number of species such as EPMI, COER, PELA  
- they also show that for PILI that among individuals that reproduce, the "potential" RA decreases with increasing propH  
- when all data points are included, all species except PEPU, PHPH, and PULU show increasing RA with increasing height  
- when RA =0 is excluded, the following species still have a significant positive correlation: BOLE, COER, EPMI, GRBU, GRSP, HATE. LEES and PELA have a p-value between 0.05-0.10   
- PEPU has a negative correlation - matches Mark's suggestion that it is "hanging on"  
- There are species - such as BOLE - which have a much "cleaner" plot against age instead of PropH (or any other size measure). This indicates that these species are programmed to reproduce at a certain "rate" regardless of their size

@@

###Looking jointly at effects of RA and PropH on growth inveestment
- Model analyzed is: lm(log([[GrowthInv]]) ~  asin(sqrt([[RA]]))*asin(sqrt([[PropH]]))))  
- The canopy species have, not suprisingly, a very obvious maxH, with Growth Investment or RGR tapering markedly at the maxH is approached. In these species growth investment is poorly correlated with RA, but well correlated with height.  
- In some of the mid-height species (GRSP, BOLE, GRBU), growth investment is very poorly correlated with height, but better correlated with RA - in other words, increased RA = less growth.  
- One of the difficult features interpreting this, is that there are species with overall higher correlations and overall lower correlations, and this pattern dominates, not the relative strength  
- My current analyses still lack the ability to tease apart the most obvious pattern - because I don't actually want to show that growth and height are correlated, so much as that the   asymptoting of growth is correlated with height - i.e. the change in slope of the relationship.  
```{r, echo=FALSE, message=FALSE}
yvar <- "GrowthInv"
xvar <- "RA"
x2var <- "PropH"

data <- SummaryInd
data <- split(data, data$species)
lmfits <- lapply(data, function(x) lm(log(x[[yvar]]) ~  asin(sqrt(x[[xvar]]))*asin(sqrt(x[[x2var]]))))
results <- summarise_all_fits_2var(lmfits)

results$r.squared <- round(results$r.squared,digits=3)
results$p.value <- round(results$p.value,digits=3)
names(results) <- c("species","r,squared","p-value","n","intercept","coef_RA","coef_PropH","coef_interaction","SS_RA","SS_PropH","SS_interaction")
results
```

#Shifts in accessory costs across species and ages
##Overall results
- Accessory costs higher for bigger seeds  
- A number of species show fairly consistent declines in accessory costs with various measures of "bigger", including size, age, RA. Other species show no patterns  
- 9 (of 14) species decrease the proportion of reproductive investment in "pre-pollination" tissues as their RA increases. in other words, they get better returns on investment the more they invest.  
- Species very different in which accessory tissues they invest the most in - which reflects the differences in basic structure (i.e. Banksia cones), but also more subtlely that some species abort lots of buds (pre-pollination aborted), while other species abort more fully formed flowers  
- For most of the statistical analyses, the youngest site has been removed, since it included individuals that were harvested while flowering (and therefore never "got to" set fruit) and automatically have no "successful" reproduction. There are of course some individuals like this at other ages as well, but far fewer

###How do accessory costs shift with seed size?
- Larger-seeded species have higher accessory costs. 
- Also shown in other studies and suggested as a mechanism for why angiosperms were able to have such small seeds (relative to gymnosperms, with no decrease in proportional accessory costs - with decreasing size)  
```{r, echo=FALSE,message=FALSE}
yvar <- "PerAcc_mean"
xvar <- "seed_size"
data <- SummarySpp

plot_yvar_vs_xvar(data, yvar, xvar,log="x",
                  xlab=xvar, ylab=yvar)
mod <- lm(log(seed_size)~PerAcc_mean,SummarySpp)
summary(mod)
anova(mod)
```
###Size versus proportion propagule
- Species that are bigger (by weight) invest a lower proportion of their reproduction in propagules  
- This is a much stronger relationship than shifts in RA by plant size  
- There is also no relationship between RA and proportion of energy invested in propagules  
- The model including both PerPropagule and RA confirms this - there is a strong species effect and strong PerPropagule effect, but RA not significant

* This means: The larger species get less of a return on their investment in reproduction. They have higher accessory costs.

- Species with larger seeds also have higher accessory costs - what we don't know is if "seed size" or "plant size" that is most important

```{r, echo=FALSE, message=FALSE}
plot(PerPropagule ~FinalWeight, data=subset(SummaryInd,RA>00),log="xy", col=col.spp[as.factor(species)],pch=16)
legend("bottomleft",legend=names(labels.spp),col=col.spp,pch=16, cex=.8,bty="n")

mod <- lm(log(FinalWeight) ~ asin(sqrt(PerPropagule))*species,data=subset(SummaryInd,ReproInv>0))
summary(mod)
anova(mod)

plot(PerPropagule_mean ~FinalWeight_mean, data=subset(SummarySppAge,RA_mean>0),log="xy", col=col.spp[as.factor(species)],pch=16)
legend("bottomleft",legend=names(labels.spp),col=col.spp,pch=16, cex=.8,bty="n")

mod <- lm(log(FinalWeight_mean) ~ asin(sqrt(PerPropagule_mean))*species,data=subset(SummarySppAge,RA_mean>0))
#summary(mod)
anova(mod)

mod <- lm(log(FinalWeight) ~asin(sqrt(PerPropagule))*asin(sqrt(RA))*species,data=subset(SummaryInd,ReproInv>0))
anova(mod)
```

### Shifts in PerPropagule and RA within species
- The above models indicate that there is no overall "PerPropagule" by "species" interaction. Looking at individual species also indicates few species show a pattern - those that do show an increase in propagule investment (decreasing accessory costs) with increasing size
    - COER, HATE, and PILI show a significant decrease in accessory costs with increasing size (p<0.05)  
    - GRSP shows a marginally significant decrease in accessory costs with increasing size (0.05<p<0.10)  
    - PEPU shows a marginally significant INCREASE in accessory costs with increasing size (0.05<p<0.10)  
    - the plots are in the file with all accessory tissues results

- There is of course also an increase in RA with increased plant weight (see RA section)
- An added piece of this, is that WITHIN species, there is an INCREASE in allocation to propagules with increasing size (age) for many species
    - BAER, COER, GRBU, GRSP, HATE, LEES, PELA, PEPU show a significant decrease in accessory costs with increasing RA (p<0.05)  
    - BOLE shows a marginally significant decrease in accessory costs with increasing RA (p ~ 0.10)

```{r, echo=FALSE, message=FALSE}
yvar <- "PerPropagule"
xvar <- "FinalWeight"

data <- filter(SummaryInd,Total>0)
data <- split(data, data$species)
lmfits <- lapply(data, function(x) lm(asin(sqrt(x[[yvar]])) ~  log(x[[xvar]])))
results <- summarise_all_fits(lmfits)

yvar <- "PerPropagule"
xvar <- "RA"

data <- filter(SummaryInd,Total>0&age>2)
data <- split(data, data$species)
lmfits <- lapply(data, function(x) lm(asin(sqrt(x[[yvar]])) ~  asin(sqrt(x[[xvar]]))))
results <- summarise_all_fits(lmfits)

results
```

###Do species differ in the proportional energy allocated to propagules (vs. accessory tissues)  
- COER, EPMI, LEES, PILI invest the most energy in propagules  
- HATE invests by far the least into propagules

```{r, echo=FALSE, message=FALSE}
yvar <- "PerPropagule"
xvar <- "species"
data <- filter(SummaryInd,Total>0)

mod <- lm(asin(sqrt(PerPropagule)) ~ species-1, data)
summary(mod)
data.frame(
    select(glance(mod), r.squared, p.value),
    n= length(resid(mod)),
    a = coef(mod)[1],
    b = coef(mod)[2]
  )

tmp <- summary(mod)
tmp <- tmp$coefficients
tmp <-as.data.frame(tmp)
tmp <- cbind(names(labels.spp),tmp)
names(tmp) <- c("species","estimate.trans","se.trans","t.value","p.value")
tmp2 <- select(SummarySpp, species, PerPropagule_mean, PerPropagule_se)
tmp <- merge(tmp, tmp2, by="species")
tmp
plot(PerPropagule_mean~species,tmp,cex.axis=.8,pch=16,log="y",ylim=c(.001,.25),main="Percent reproductive investment put into propagules")

mean1 <- tmp$PerPropagule_mean
se1 <- tmp$PerPropagule_se

se_function(mean1,se1)
```
